\thetitle

### 2021-04-29 paper/paper.tex
update

\MS[t]{Algorithm adaptions}

\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete.
%
% Recall that we gproducts routed over the bottleneck machine \(M5\) are bottleneck products.
%
For the \textit{state space} we adapt the result of \citet{knollmann2013control}, which state that
the lead times ought to control the WIP level, instead of the due date reliability as changes are
directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
information for each product type \(\pt\):
\begin{itemize}

@@

% N-Step: 5
\textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
replay memory, and connects the state values starting from the latest observed experience, such that
the newly computed state value of the corresponding experience propagates to the experience of the
previous period. This allows a faster convergence, as values bootstrap over multiple periods in
contrast to just one step. The experiences are stored according to the first taken action in the
n-step series. We uniformly sample over the actions to ensure each action is constantly trained.
This adaption reduces catastrophic forgetting of already accumulated knowledge.
%
% Overestimating
% The idea behind overestimating the average reward is to constantly seek for better policies than the
% currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
% \(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
% values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
% overestimation equals \(0\).
% Rho Minimum
% Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
% and provides a methodology for constantly improving policy.
%
% Workers Min Exploration [0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20]
Furthermore, we use \(8\) additional \textsf{worker threads} (parallel RL) that operate on the same
environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.

@@

\paragraph*{Neural Network Setup.}
%
Due to the vast possibilities the neural network architecture was manually optimised, starting with
less stochastic systems and aiming for a coffin shape that can be used in all setups. During this
process we constantly adapted and reevaluated the best network on various production system
configurations. The resulting architecture depends on the number of features \(\anninp\) and the
number of output nodes \(\annout\), which itself depends on the number of actions
\(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
Q-Learning only approximates the function \(Q_{\gamma}^\pol\).
%
The resulting feedforward deep ANN architecture is given as %
\(\anninp %
\rightarrow 1.75 \anninp % \rightarrow \relu
\rightarrow 1.5 \anninp % \rightarrow \relu
% \rightarrow 2 \annout % \rightarrow \relu
\rightarrow \annout %
% \rightarrow   \leakyTanh%
\), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\). Note
that we explicitly do not use an output activation function. The idea is that we do not want to
restrict the approximations to values inside this range. This is especially important for
Q-Learning, as the range of the state values are difficult to forecast.

The values \(\anninp\) and \(\annout\) depend on the feature extraction method. In our case case
using 6 independent agents we gain \MS[t]{todo: check} \(\anninp=540\) and \(\annout=36\)
\MS[t]{check setup}. For the Q-Learning algorithm, which only approximates one function, the output
nodes \(\annout\) are halved. \MS[t]{what is the correct size?}

% Min-Max Scaling of input/output. Output: ScalingNetOutParameters (-800) 800 (-300) 300 (-400) 800 (-400) 800
\MS[t]{scaling of input layer} The features representing order counts are scaled using min-max
scaling with minimum \(0\) and maximum \(5\), the lead times with minimum \(0\) and maximum \(7\)
respectively.
% Output layer scaling
To further ensure smaller ANN parameter values the output is transformed using min-max scaling to
the interval \((-400, 800)\), s.t.\@ an ANN output of \(1\) represents a value of \(800\).
%
% Training Mini-Batch Size:  4
On each training step, i.e.\@ every n-step steps, randomly \(m\) sets of experiences the training
mini-batch size each consisting of n-step experiences are selected, there gradients computed and the
sum of gradients back-propagated though the worker network using Adam \citep{kingma2014adam}, where
the weight decay with parameter \(\lambda\) is implemented as presented by
\citet{loshchilov2017decoupled}.
%
% Smooth target network period: 100 with update 0.01
Every \(C\) steps the target network is updated by exponentially smoothing the parameters with rate
\(\gamma\).
%
% ANN Init
As neural network
initialisation method we use the commonly known Xavier initialisation
\citep{glorot2010understanding}, i.e.\@ the initial value for each weight \(i\) is sampled from a
uniform distribution \(W_{l,i} \sim \Unif(-1/\sqrt{n_l},1/\sqrt{n_l})\), where \(n_{l}\) is the
number of nodes in layer \(l\).
%
%

@@

\MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
\cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
the production system. That is, although orders are immediately released to the production system
and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
takes several periods to materialise. For instance, by deciding not to release an order, which later
becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
instead of the last action (only). Thus they propose to keep track of the orders currently in the
order pool until all fully traversed through the production system and reward the actions according
to all orders in the order pool.

\begin{table}[t]
  \centering
  \footnotesize
  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
    \toprule
    Parameter                      & $\alpha$   & $\alpha_{\max}$ & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
    \midrule
    Start value                    & 0.01       & \(2^{-5}\)      & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
    Exp. Smth. Rate                & 0.25       &                 &              &              & 0.25     &            & 0.25      &           \\
    Exp. Smth. Steps in \(10^{3}\) & 50         &                 &              &              & 100      &            & 100       &           \\
    Minimum value                  & \(5^{-5}\) &                 &              &              & 0.001         &            & 0.005     &           \\
    \bottomrule
    \toprule
    Parameter                      & \multicolumn{8}{c}{Value} \\
    \midrule
    Workers Min. Exploration & \multicolumn{8}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}\\
    Training Batch Size & \multicolumn{8}{c}{4}\\
    Target Network Update & \multicolumn{8}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}\\
    N-Step & \multicolumn{8}{c}{5}\\
    Training Mini-Batch Size & \multicolumn{8}{c}{\(m = 4\)}\\
    Adam Back-Propagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999 1e-8 1e-3,
    \bottomrule
  \end{tabular}
  \caption{\label{tbl:params}Parameter Setup.}
\end{table}

The parameter setup is given in Table~\ref{tbl:params}. The parameters are exponentially decayed
with rate \(0.05\) and \(350k\) steps\footnote{The decay is defined as \(0.05 ^{(t/350k)}\), where
  \(t\) is the current period}. \MS[t]{TODO} We implemented the deep Q-network algorithm proposed by
\cite{mnih2015human},  a discount factor of
\(0.99\).

@@

>>>>

\paragraph{Markov Decision Process.} The underlying MDP is unichain and looks as follows. Both,
state space $\mathcal{S}$ and action space $\mathcal{A}$ are discrete.
%
% Recall that we products routed over the bottleneck machine \(M5\) are bottleneck products.
%
For the \textit{state space} we adapt the result of \citet{knollmann2013control}, which state that
the lead times ought to control the WIP level, instead of the due date reliability as changes are
directly seen. Thus, any state $s \in \mathcal{S}$ of the state space is composed of the following
information for each product type \(\pt\):
\begin{itemize}

@@

% N-Step: 5
As in \citep{mnih2016asynchronous, mnih2015human} the algorithm uses a \textit{Replay Memory} to
save experiences and then randomly selects experiences from there for learning. This ensures that
the independent and identically distributed (i.i.d.) assumption of the samples for the ANN
backpropagation algorithms are satisfied. We extend this idea by splitting the available size of the
replay memory into subsets, where each subset represents one action and sample uniformly over these
subsets first. This prevents overfitting when one action is overrepresented in the policy, e.g.\@ no
lead time change occurs much more often than the other actions.

\textit{N-step learning} collects a set of consecutive experiences, in our case retrieved from the
replay memory, and connects the state values starting from the latest observed experience, such that
the newly computed state value of the corresponding experience propagates to the experience of the
previous period. This allows a faster convergence, as values bootstrap over multiple periods in
contrast to just one step. The experiences are stored according to the first taken action in the
n-step series. We uniformly sample over the actions to ensure each action is constantly trained.
This adaption reduces catastrophic forgetting of already accumulated knowledge.
%
% Overestimating
% The idea behind overestimating the average reward is to constantly seek for better policies than the
% currently known. We overestimate the reward by subtracting \(10\%\) of the computed average reward
% \(\avgrew^{\pol}\) and an exponentially smoothed average reward of the actually returned reward
% values of greedy actions. Thus, if the computed and actually observed average reward coincide, the
% overestimation equals \(0\).
% Rho Minimum
% Preventing an increase of the average reward with \(\avgrew_{\max}^{\pol}\) reduces the search space
% and provides a methodology for constantly improving policy.
%
% Workers Min Exploration [0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20]
Furthermore, we use \(8\) additional \textsf{worker threads} (parallel RL) that operate on the same
environment to collect a broader range of experiences~\citep[as e.g. in][]{mnih2016asynchronous}.

@@

\paragraph*{Neural Network Setup.}
%
Due to the vast possibilities the neural network architecture was manually optimised, starting with
less stochastic systems and aiming for a coffin shape that can be used in all setups. During this
process we constantly adapted and reevaluated the best network on various production system
configurations. The resulting architecture depends on the number of features \(\anninp\) and the
number of output nodes \(\annout\), which itself depends on the number of actions
\(\acts = \size{\States}\), the independent agent configuration, and the number of functions that
are approximated. For the later, in case of \ARA{} this is \(2\), as we represent both functions
\(\X^\pol_{\gamma_{0}}\) and \(\X^\pol_{\gamma_{1}}\) with the same neural network, while deep
Q-Learning only approximates the function \(Q_{\gamma}^\pol\).
%
The resulting feedforward deep ANN architecture is given as %
\(\anninp %
\rightarrow 1.75 \anninp % \rightarrow \relu
\rightarrow 1.5 \anninp % \rightarrow \relu
% \rightarrow 2 \annout % \rightarrow \relu
\rightarrow \annout %
% \rightarrow   \leakyTanh%
\), where each arrow symbolises a fully-connected layer with a leaky rectified linear unit
(\(\relu\)) as activation function, which leaks values smaller than \(0\) with rate \(0.02\).
%
Note
that we explicitly do not use an output activation function. The idea is that we do not want to
restrict the approximations to values inside this range. This is especially important for
Q-Learning, as the range of the state values are difficult to forecast.
%
The actual values \(\anninp\) and \(\annout\) depend on the used algorithm. In our case case using 6
independent agents we gain \(\anninp=324\) and \(\annout=36\). For the Q-Learning algorithm, which
only approximates one function, the output nodes \(\annout\) are halved.
%
% SpecFullyConnected 324 567 :=> SpecLeakyRelu (567,1,1) :=> SpecFullyConnected 567 486 :=>
% SpecLeakyRelu (486,1,1) :=> SpecFullyConnected 486 36 :=> SpecReshape (36,1,1) (18,2,1) :=>
% SpecNNil2D18x2
%

% Min-Max Scaling of input/output. Output: ScalingNetOutParameters (-800) 800 (-300) 300 (-400) 800 (-400) 800
The features representing order counts are scaled using min-max scaling with minimum \(0\) and
maximum \(5\), the lead times with minimum \(0\) and maximum \(7\) respectively.
% Output layer scaling
To further ensure smaller ANN parameter values the output is transformed using min-max scaling to
the interval \((-400, 800)\), s.t.\@ an ANN output of \(1\) represents a value of \(800\).
%
% Training Mini-Batch Size:  4
On each training step, i.e.\@ every n-step steps, randomly \(m\) sets of experiences the training
mini-batch size each consisting of n-step experiences are selected, there gradients computed and the
sum of gradients backpropagated though the worker network using Adam \citep{kingma2014adam}, where
the weight decay with parameter \(\lambda\) is implemented as presented by
\citet{loshchilov2017decoupled}.
%
% Smooth target network period: 100 with update 0.01
Every \(C\) steps the target network is updated by exponentially smoothing the parameters with rate
\(\gamma\).
%
% ANN Init
As neural network
initialisation method we use the commonly known Xavier initialisation
\citep{glorot2010understanding}, i.e.\@ the initial value for each weight \(i\) is sampled from a
uniform distribution \(W_{l,i} \sim \Unif(-1/\sqrt{n_l},1/\sqrt{n_l})\), where \(n_{l}\) is the
number of nodes in layer \(l\).
%
%

@@

% \MS[t]{todo} We evaluate two different kinds of reward reporting as they have been proposed by
% \cite{Schneckenreither2019}. They find that it is beneficial to overcome the time offset imposed by
% the production system. That is, although orders are immediately released to the production system
% and thus generate a response in terms of a reward to the agent, the full impact of the chosen action
% takes several periods to materialise. For instance, by deciding not to release an order, which later
% becomes a backorder, it is beneficial to reward the actions actually responsible for the backorder
% instead of the last action (only). Thus they propose to keep track of the orders currently in the
% order pool until all fully traversed through the production system and reward the actions according
% to all orders in the order pool.

\begin{table}[t]
  \centering
  \footnotesize
  \begin{tabular}{lP{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}P{1cm}}
    \toprule
    Parameter                      & $\alpha$   & $\alpha_{\max}$ & $\gamma_{0}$ & $\gamma_{1}$ & $\gamma$ & $\epsilon$ & $p_{exp}$ & $\plearn$ \\
    \midrule
    Start value                    & 0.01       & \(2^{-5}\)      & 0.8          & 0.99/1.0     & 0.01     & 0.25       & 1.0       & 0.5       \\
    Exp. Smth. Rate                & 0.25       &                 &              &              & 0.25     &            & 0.25      &           \\
    Exp. Smth. Steps in \(10^{3}\) & 50         &                 &              &              & 100      &            & 100       &           \\
    Minimum value                  & \(5^{-5}\) &                 &              &              & 0.001         &            & 0.005     &           \\
    \bottomrule
    \toprule
    Parameter                      & \multicolumn{8}{c}{Value} \\
    \midrule
    Workers Min. Exploration & \multicolumn{8}{c}{0.01, 0.02, 0.03, 0.04, 0.05, 0.10, 0.15, 0.20}\\
    Training Batch Size & \multicolumn{8}{c}{4}\\
    Target Network Update & \multicolumn{8}{c}{Exp. smooth. with rate \(\gamma\) every \(C=100\) steps}\\
    N-Step & \multicolumn{8}{c}{5}\\
    Training Mini-Batch Size & \multicolumn{8}{c}{\(m = 4\)}\\
    Adam ANN Backpropagation & \multicolumn{8}{c}{\(\alpha = 0.005, \beta_{1}=0.9,
                       \beta_{2}=0.999, \epsilon=10^{-8}, \lambda=0.001}\\ % OptAdam 0.005 0.9 0.999
% 1e-8 1e-3,
        Replay Memory Size & \multicolumn{8}{c}{\(10800\)}\\
    \bottomrule
  \end{tabular}
  \caption{\label{tbl:params}Overview of Parameter Setup.}
\end{table}

Table~\ref{tbl:params} presents an overview of the parameter setup. The parameters \(\alpha\),
\(\gamma\) and \(p_{exp}\) are exponentially decayed as indicated.% \footnote{The decay is defined as
  % \(0.05 ^{(t/350k)}\), where \(t\) is the current period}.
\MS[t]{TODO} As further comparison we implement the deep Q-Learning network algorithm proposed
by~\cite{mnih2015human}, where for comparison reasons we also integrate n-Step learning, the
experience replay memory adaptions and independent agents. \MS[t]{Discount factor in table?}

% a discount factor of \(0.99\).

@@

\MS[t]{Multiple agents on different tasks (hierarchical PP): e.g. Order acceptance + release,
  scheduling + release, etc.}
